Supporting Information

### 2018-03-03 SupportingInformation.Rmd
additions to copula functions and application to C/N data

<!--Setup-->
```{r setup, echo=F}
knitr::opts_chunk$set(echo = TRUE, fig.pos = "H")

@@

seed<-101
families<-c(1,3:10,13,14,16:20)
#A function needed for caching
mtime <- function(files)
{
  lapply(Sys.glob(files),function(x) X =   
          file.info(x)$mtime)
}
source("getcopula.R")
source("OurBiCopSelect.R")
source("FittingCopula_selective_loc.R")
source("aphid_data_calling.R")
source("plankton_north_sea_data_calling.R")
source("NonParamStat.R")
source("copsurrog2d.R")
```

@@

## Soil C and N data
Carbon and nitrogen in soil at many locations across CONUS (Terry to augment this description).

@@

<!--The copula looks like this:-->
```{r rankandplot_Loecke, echo=F, fig.cap="Copula plot of soil C and N data",  fig.height=5, fig.width=5}
#convert to a copula and plot
fig_nums("soilCNcop",display=F)
v<-getcopula(d=d,rankon=T,ploton=T) #with real data, rankon should be T
```

The raw data looks like : `r fig_nums("soilCNraw",display="c")`.

The copula looks like : `r fig_nums("soilCNcop",display="c")`.

\newpage
## Aphid data
   * Aphid data contains 3 measurements [count, firstflight, flight-duration] at 11 locations for 20 species and for 35 years.
   * We threw out all locations that had fewer than 30 years data 
   * We use species=10 [Name : "Green spruce aphid"] for aphid-count data 
   * We use species=11 [Name : "Leaf-curling plum aphid"] for aphid-firstflight data

@@

<!--Shya pls add a description of each of the statistics implemented in CopulaFunctions.R. As usual, please describe the stats with sufficient precision that they could be re-coded by the reader.-->
#### stat_CorlCoru
```{r stat_CorlCoru_D2lD2u, echo=F}
fig_nums("stat_CorlCoru_D2lD2u",display=F)
```
![schematic diagram for stat_CorlCoru and stat_D2lD2u](./CorlCoru_D2lD2u.jpg){ width=30%, height=30% }

@@

### Model selection approach
```{r bivmodselect_Loecke, echo=F, cache=T, cache.extra=list(seed,v,mtime("OurBiCopSelect.R"))}
set.seed(seed)
families_Loecke<-c(1,3:10,13:14,16:20) 
BivMS_res_Loecke<-OurBiCopSelect(u1=v[,1],u2=v[,2],families=families_Loecke,level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,gofnormal=F,status=F)
```
The test of indepedence gave $p=$ `r BivMS_res_Loecke$IndepTestRes`.

@@

### Non-paramtric analysis of tail dependence
<!--Do the intensive part of the analysis-->
```{r stat_soilCN, echo=F, cache=T, cache.extra=list(seed,v,mtime("CopulaFunctions.R"),mtime("copsurrog2d.R"),mtime("NonParamStat.R"))}
set.seed(seed) 
resloc<-"./Results/stat_results/stat_soilCN/"

#get the stats for the real data
stat_soilCN<-copsync(v) 
saveRDS(stat_soilCN,paste(resloc,file="stat_soilCN.RDS",sep=''))
sCN<-stat_soilCN[c(2:5,7,8,10:13)]
stat<-unlist(sCN)
stat<-c(stat[1:4],unname(stat[3]-stat[4]),stat[5:6],unname(stat[5]-stat[6]),stat[7:8],unname(stat[7]-stat[8]),stat[9:10],unname(stat[10]-stat[9]))
names(stat)[c(5,8,11,14)]<-c("Corl-Coru","Pl-Pu","Tl-Tu","D2u-D2l")

#get surrogates and get stats for them, preserving kendall
cop<-normalCopula(.5,2)
numsurrog<-1000
surv<-copsurrog2d(v,cop,"kendall",numsurrog)
statsurr_k<-matrix(NA,length(stat),numsurrog)
rownames(statsurr_k)<-names(stat)
for (counter in 1:dim(surv)[3])
{
  h<-copsync(surv[,,counter])
  h<-unlist(h[c(2:5,7,8,10:13)])
  h<-c(h[1:4],h[3]-h[4],h[5:6],h[5]-h[6],h[7:8],h[7]-h[8],h[9:10],h[10]-h[9])
  statsurr_k[,counter]<-unname(h)
}

#same, preserving spearman
surv<-copsurrog2d(v,cop,"spearman",numsurrog)
statsurr_s<-matrix(NA,length(stat),numsurrog)
rownames(statsurr_s)<-names(stat)
for (counter in 1:dim(surv)[3])
{
  h<-copsync(surv[,,counter])
  h<-unlist(h[c(2:5,7,8,10:13)])
  h<-c(h[1:4],h[3]-h[4],h[5:6],h[5]-h[6],h[7:8],h[7]-h[8],h[9:10],h[10]-h[9])
  statsurr_s[,counter]<-unname(h)
}
```

<!--Display the table generated above-->
```{r table_stat_soilCN, echo=F, results="asis"}
#now get fractions of surrogates lt the data and gt the data for each stat
stattab<-data.frame(statname=names(stat),stat=unname(stat),frac_surr_gt_k=NA,frac_surr_lt_k=NA,frac_surr_gt_s=NA,frac_surr_lt_s=NA)
for (counter in 1:length(stat))
{
  stattab[counter,3]<-sum(statsurr_k[counter,]>stat[counter])/numsurrog
  stattab[counter,4]<-sum(statsurr_k[counter,]<stat[counter])/numsurrog
  stattab[counter,5]<-sum(statsurr_s[counter,]>stat[counter])/numsurrog
  stattab[counter,6]<-sum(statsurr_s[counter,]<stat[counter])/numsurrog
}

#display
names(stattab)<-c("Statistic","Value on data","Frac_K_sur_gt","Frac_K_sur_lt","Frac_S_sur_gt","Frac_S_sur_lt")
table_nums("table_stat_soilCN",display=F)
knitr::kable(stattab,caption = "Results of non-parametric tests of tail dependence and asymmetry of tail dependence for the soil C and N dataset. Names and values on the actual C and N data are listed in the first two columns. The third (respectively, fourth) column is the fraction of Kendall-preserving surrogate datasets for which the given statistic is greater than (resp., less than) its value on the real data. The last two columns are analogous but for Spearman-preserving surrogates.")
```

>>>>

<!--Basic setup-->
```{r setup, echo=F}
knitr::opts_chunk$set(echo = TRUE, fig.pos = "H")

@@

seed<-101
families<-c(1,3:10,13,14,16:20)

#A function needed for caching
mtime <- function(files)
{
  lapply(Sys.glob(files),function(x) X =   
          file.info(x)$mtime)
}

source("getcopula.R")
source("OurBiCopSelect.R")
source("FittingCopula_selective_loc.R")
source("aphid_data_calling.R")
source("plankton_north_sea_data_calling.R")
source("NonParamStat.R")
source("copsurrog2d.R")
```

@@

@@

<!--The copula looks like this:-->
```{r rankandplot_Loecke, echo=F, fig.cap="Copula plot of soil C and N data",  fig.height=5, fig.width=5}
#convert to a copula and plot
fig_nums("soilCNcop",display=F)
v_CN<-getcopula(d=d,rankon=T,ploton=T) #with real data, rankon should be T
```

Carbon and nitrogen in soil at many locations across CONUS (Terry to augment this description). The raw data are shown in `r fig_nums("soilCNraw",display="c")`. The copula is shown in `r fig_nums("soilCNcop",display="c")`.

<!--Shyamolina, let's not put in newpages and other spacing modifications until possibly at the end, once all the 
content is essentially settled-->

## Aphid data
   * Aphid data contains 3 measurements [count, firstflight, flight-duration] at 11 locations for 20 species and for 35 years.
   * We threw out all locations that had fewer than 30 years data 
   * We use species=10 [Name : "Green spruce aphid"] for aphid-count data 
   * We use species=11 [Name : "Leaf-curling plum aphid"] for aphid-firstflight data

@@

#### stat_CorlCoru
```{r stat_CorlCoru_D2lD2u, echo=F}
fig_nums("stat_CorlCoru_D2lD2u",display=F)
```
![schematic diagram for stat_CorlCoru and stat_D2lD2u](./CorlCoru_D2lD2u.jpg){ width=30%, height=30% }

@@

### Model selection approach
```{r bivmodselect_Loecke, echo=F, cache=T, cache.extra=list(seed,v_CN,mtime("OurBiCopSelect.R"))}
set.seed(seed)
families_Loecke<-c(1,3:10,13:14,16:20) 
BivMS_res_Loecke<-OurBiCopSelect(u1=v_CN[,1],u2=v_CN[,2],families=families_Loecke,level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,gofnormal=F,status=F)
```
The test of indepedence gave $p=$ `r BivMS_res_Loecke$IndepTestRes`.

@@

### Non-paramtric analysis of tail dependence
<!--Do the intensive part of the analysis-->
```{r stat_soilCN, echo=F, cache=T, cache.extra=list(seed,v_CN,mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
set.seed(seed)

#initialize
npa_res_CN<-data.frame(Statistic=c("Kendall","Spearman","Corl","Coru","CorlmCoru","Corlt0p1","Corgt0p9","Corlt0p1mCorgt0p9","D2l","D2u","D2umD2l","D2lt0p1","D2gt0p9","D2lt0p9mD2gt0p1"),Value=NA,Frac_of_K_surr_gt_val=NA,Frac_of_S_surr_gt_val=NA)

#utility function
calcstats<-function(v)
{
  h1<-Corbds(v[,1],v[,2],0,.5)
  h2<-Corbds(v[,1],v[,2],0.5,1)
  h3<-Corbds(v[,1],v[,2],0,.1)
  h4<-Corbds(v[,1],v[,2],0.9,1)
  h5<-D2bds(v[,1],v[,2],0,.5)
  h6<-D2bds(v[,1],v[,2],0.5,1)
  h7<-D2bds(v[,1],v[,2],0,.1)
  h8<-D2bds(v[,1],v[,2],0.9,1)
  return(c(cor(v[,1],v[,2],method="kendall"),
           cor(v[,1],v[,2],method="spearman"),
           h1,h2,h1-h2,
           h3,h4,h3-h4,
           h5,h6,h6-h5,
           h7,h8,h8-h7
  ))
}

#get value on the real data
npa_res_CN[,"Value"]<-calcstats(v_CN)

#Do the kendall surrogates
cop<-normalCopula(.5,2)
numsurrog<-1000
surv<-copsurrog2d(v_CN,cop,"kendall",numsurrog)
res<-apply(FUN=calcstats,X=surv,MARGIN=3)
for (counter in 1:dim(npa_res_CN)[1])
{
  npa_res_CN[counter,"Frac_of_K_surr_gt_val"]<-sum(res[counter,]>npa_res_CN[counter,"Value"])/numsurrog
}

#Do the spearman surrogates
surv<-copsurrog2d(v_CN,cop,"spearman",numsurrog)
res<-apply(FUN=calcstats,X=surv,MARGIN=3)
for (counter in 1:dim(npa_res_CN)[1])
{
  npa_res_CN[counter,"Frac_of_S_surr_gt_val"]<-sum(res[counter,]>npa_res_CN[counter,"Value"])/numsurrog
}
```

<!--Display the table generated above-->
```{r table_stat_soilCN, echo=F, results="asis"}
table_nums("table_stat_soilCN",display=F)
knitr::kable(npa_res_CN,caption = "Results of non-parametric tests of tail dependence and asymmetry of tail dependence for the soil C and N dataset. Names of statistics and values on the actual C and N data are listed in the first two columns. The third (respectively, fourth) column is the fraction of Kendall-preserving (respectively, Spearman-preserving) surrogate datasets for which the given statistic is greater than its value on the real data. So if these quantities are smaller than 0.025 the statistic is significantly larger than expected based on null hypotheses of normal copulas; if larger than 0.975, significantly smaller.")
```
